---
title: "Untitled"
author: "Angella Nabasirye"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(dplyr)
library(ggplot2)
library(tidyr)
library(gridExtra)
library(knitr)
library(grid)
```

## Load Data
```{r cars}
df <- read.csv("D:/intro to R/R practise datasets/Maternal Health Risk Data Set.csv")
```

## Mean Age by Risk Level
```{r pressure, echo=FALSE}
mean_age_by_risk <- df %>%
  group_by(RiskLevel) %>%
  summarise(mean_age = mean(Age, na.rm = TRUE)) %>% arrange(desc(mean_age))

print(mean_age_by_risk)
```
## Relationship Between Age and Risk Level. 
- The mean age increases as the risk level increases
- Older individuals are more likely to fall into the high-risk category.
- Younger individuals tend to be in the low-risk category.
- This suggests that age is a contributing factor to health risk levels, possibly due to age-related health conditions like hypertension, diabetes, or cardiovascular diseases.

## median_diastolic_by_risk

```{r}
median_diastolic_by_risk <- df %>%
  group_by(RiskLevel) %>%
  summarise(median_diastolic = median(DiastolicBP, na.rm = TRUE)) %>%
  arrange(median_diastolic)

print(median_diastolic_by_risk)
```
## Relationship Between Diastolic Blood Pressure and Risk Level
- The median diastolic blood pressure (BP) is higher for high-risk individuals
- Individuals in the high-risk group have significantly higher diastolic BP compared to the mid- and low-risk groups.
- Diastolic BP of 90 mmHg suggests hypertension (high blood pressure), which is a strong risk factor for cardiovascular diseases.
- The low and mid-risk groups have similar median diastolic BP (75 mmHg), meaning there might not be a significant difference in BP between these two groups.

## Summary of the Output
- The first analysis tells us which risk level has the highest average age.
- The second analysis shows which risk level has the lowest and highest median diastolic blood pressure.
- These insights help in understanding how age and blood pressure correlate with risk levels.

## QN 2 Creating a new column SystolicRate
```{r}
df <- df %>% mutate(SystolicRate = SystolicBP / Age)
```
## QN 3 
```{r}
filtered_data <- df %>%
  mutate(SystolicRate = SystolicBP / Age)
```
## Filter age ≤ 40
```{r}
colnames(df)
head(df) 
str(df)
```
```{r}
filtered_data <- df %>%
  filter(Age <= 40) %>%  
  group_by(RiskLevel) %>%
  summarise(mean_systolic_rate = mean(SystolicRate, na.rm = TRUE)) %>%  
  arrange(desc(mean_systolic_rate))

print(filtered_data)
```

## Qn 4 Selecting continuous variables
```{r}
continuous_vars <- c("Age", "SystolicBP", "DiastolicBP", "BS", "BodyTemp", "HeartRate", "SystolicRate")
```

### Converted dataset to long format for easy plotting
```{r}
df_long <- df %>%
  pivot_longer(cols = all_of(continuous_vars), names_to = "Variable", values_to = "Value")
```

### Plot histograms
```{r}
ggplot(df_long, aes(x = Value)) +
  geom_histogram(fill = "skyblue", color = "black", bins = 30) +
  facet_wrap(~ Variable, scales = "free") +
  labs(title = "Histograms of Continuous Variables", x = "Value", y = "Frequency") +
  theme_minimal()
```

### Boxplots by Risk Level
```{r}

```


